### 2004-07-15 Documentation/Manual.tex
Minor changes only The spaces in index entries with \_ remains, it seems to be rooted in the way how \index processes its arguments -> we have to edit the idn-files at last by hand :-<
added the installation section in app A added targte 'fast' to Makefile
*** empty log message ***

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLB}[1]{{\tt glb#1}\index{api}{#1@{\tt #1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt #1}}}

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{{\tt  #1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNSC}[1]{\index{constants}{{\tt #1}}}
\newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
\newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}

\definecolor{light}{gray}{0.93}
\definecolor{heavy}{gray}{0.35}
\setlength{\fboxrule}{0.1cm}
\newcommand{\example}[2]{
\begin{table}[p]
\begin{center}
\fcolorbox{heavy}{light}{
\begin{minipage}[t][\textheight][t]{15.5cm}
{\em \underline{Example:} #1}

@@

\begin{center}
{ \Large \bf
\GLOBES\ 1.X \\
General Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and reference manual }

\vspace*{1.5cm}
{\large \bf !!!DRAFT IN PROGRESS!!! }
\end{center}

@@

\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino long baseline experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language, which allows to describe most classes of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for Linux. Since the source code is included,
the modifications to other operating systems should be doable.

@@

As it is illustrated in \figu{GLOBES}, \GLOBES\ consists 
of several modules.
%
\begin{figure}[bht]
\begin{center}
\includegraphics[width=16cm]{GLOBES}
\end{center}
\caption{\label{fig:GLOBES} Different modules in \GLOBES .}
\end{figure}
%
\AEDL (``Abstract Experiment Definition Language'') is a language
to define experiments in form of normal text files. One or more of 
the resulting \AEDL\ files can then be processed together with supporting 
flux or cross section files by the user interface. The user interface
is a C-library, which loads one or more \AEDL\ file(s)
containing the experiment definition(s). The user interface is linked 
against the application software, which calls the user interface functions
for the intended experiment simulation.

The application 
software is, except from some example files, not part of \GLOBES , since
the evaluation of the experiment performance is often a matter of taste
and definition. In addition, the algorithms depend, especially for
high-precision instruments, very much on the oscillation parameters.
In general, it is quite simple to simulate superbeams and reactor
experiments. However, because of the more complicated topology, the
simulation of neutrino factories is much more difficult. In order
to demonstrate some of these difficulties, we show in this manual only
examples with neutrino factories. These examples can be found in
\Part~\ref{part:1} within the boxed pages. As complete files, they
are also available in the \GLOBES\ software package.

@@

The user interface functions are described in \Part~\ref{part:1} of 
this manual, which is the ``user's manual''. In there, first of all a 
short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
have an overview over \GLOBES . 
After that, the user
interface is successively introduced from very basic to more sophisticated
function. Eventually, it is demonstrated how one can change many
experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
information. We recommend that everybody interested in \GLOBES\ should
become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
 and some of the examples on the boxed pages. The examples can be 
 directly compiled 
 from the respective directory in the \GLOBES\ software package.

In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
introductory chapter, all functions are defined in greater detail.
This part might be more interesting for the experimental users who
want to modify or create \AEDL\ files. A useful tool in this context
is the software program ``globes'', which returns event rates and other
information for individual \AEDL\ files without further programming. 
For example, flux normalizations can with this tool be easily adjusted 
to reproduce the event rates of a specific experiment.

>>>>

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLB}[1]{{\tt glb#1}\index{api}{#1@{\tt #1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt \protect{#1}}}}

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\protect{\tt{#1}}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNSC}[1]{\index{constants}{{\protect{\tt #1}}}}
\newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
\newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}
\newcommand{\bugs}{{\tt globes@ph.tum.de}}
\definecolor{light}{gray}{0.93}
\definecolor{heavy}{gray}{0.35}
\setlength{\fboxrule}{0.1cm}
\newcommand{\example}[2]{
\begin{table}[p]
\begin{center}
\fcolorbox{heavy}{light}{
\begin{minipage}[t][\textheight][t]{15.5cm}
{\em \underline{Example:} #1}

@@

\begin{center}
{ \Large \bf
\GLOBES\ 1.X \\
General Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and experiment definition manual }

%\vspace*{1.5cm}
%{\large \bf !!!DRAFT IN PROGRESS!!! }
\end{center}

@@

\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language, which allows to describe most classes of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for Linux. Since the source code is included,
the modifications to other operating systems should be doable.

@@

As it is illustrated in \figu{GLOBES}, \GLOBES\ consists 
of several modules.
%
\begin{figure}[bht]
\begin{center}
\includegraphics[width=16cm]{GLOBES}
\end{center}
\caption{\label{fig:GLOBES} Different modules in \GLOBES .}
\end{figure}
%
\AEDL (``Abstract Experiment Definition Language'') is a language
to define experiments in form of ordinary text files. One or more of 
the resulting \AEDL\ files can then be processed together with supporting 
flux or cross section files by the user interface. The user interface
is a C-library, which loads one or more \AEDL\ file(s)
containing the experiment definition(s). The user interface is linked 
against the application software, and provides the user interface functions
for the intended experiment simulation.

The application 
software is, except from some example files, not part of \GLOBES , since
the evaluation of the experiment performance is often a matter of taste
and definition. In addition, the algorithms depend, especially for
high-precision instruments, very much on the oscillation parameters.
In general, it is quite simple to simulate superbeams and reactor
experiments. However, because of the more complicated topology, the
simulation of neutrino factories is much more difficult. In order
to demonstrate some of these difficulties, we present in this manual only
examples with neutrino factories. These examples can be found in
\Part~\ref{part:1} within the boxed pages. As complete files, they
are also available in the \GLOBES\ software package.

@@

The user interface functions are described in \Part~\ref{part:1} of 
this manual, which is the ``user's manual''. In there, first of all a 
short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
have an overview over \GLOBES . 
After that, the user
interface is successively introduced from very basic to more sophisticated
functions. Eventually, it is demonstrated how one can change many
experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
information. We recommend that everybody interested in \GLOBES\ should
become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
 and some of the examples on the boxed pages. The examples can be 
 directly compiled 
 from the respective directory in the \GLOBES\ software package.

In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
introductory chapter, all functions are defined in greater detail.
This part might be more interesting for the experimental users who
want to modify or create \AEDL\ files. A useful tool in this context
is the software program ``globes'', which returns event rates and other
information for individual \AEDL\ files without further programming. 
For example, flux normalizations can with this tool be easily adjusted 
to reproduce the event rates of a specific experiment. It is described
in the last chapter of \Part~\ref{part:2}.

### 2004-07-15 Documentation/appendix.tex
*** empty log message ***

The installation of \GLOBES\ is highly automated and there should not be
any problems on a decently up-to-date GNU/Linux system. The installation
has however only been tested on a limited number of platforms, mainly
running with various versions of SuSE Linux. Thus we would appreciate
to know your experiences with installation on different platforms, please
send e-mail to <\bugs>.

\section*{Prerequisites for installation of \GLOBES}
\index{norm}{Installation!prerequisites}

Besides the usual things like a working libc you need to have
\begin{itemize}
\item[gcc] The GNU compiler collection\\
\verb+gcc.gnu.org+
\item[BLAS] Basic Linear Algebra Subprograms\\
\verb+www.netlib.org/blas/+    
\item[LAPACK] Linear Algebra PACKage\\
\verb+www.netlib.org/lapack/+
\item[f2c] Fortran to C\\ 
\verb+www.netlib.org/f2c/+
\item[GSL] The GNU Scientific Library\\
\verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any
ANSI C/C++ compiler but the \verb+globes+ binary uses the 
\verb+argp+ facility of \verb+glibc+ to parse its command line options.

All those libraries are also available as rpm's from the various
distributors of GNU/Linux, see their web sites for downloads. Due to
the rather difficult nature of the build process of BLAS and LAPACK
it is recommended to use a suitable rpm instead. Chances are that
gcc, f2c and GSL are already part of your installation. Furthmore you need
a working \verb+make+ to build and install \GLOBES.

In future releases it is planed to at least get rid of LAPACK, because
then also BLAS and f2c would drop out. Furthemore there is no prinicipal
reason why not also to get rid of any C++ parts. This will greatly simplify
the build process and ease the requirements for installation.

@@

In general it is preferable to use the shared libraries, in case you don't
have root privileges see the corresponding section of this file.

@@

   The \verb+configure+ shell script attempts to guess correct values for
various system-dependent variables used during compilation.  It uses
those values to create a \verb+Makefile+ in each directory of the package.
Finally, it creates a shell script \verb+config.status+ that
you can run in the future to recreate the current configuration, a file
\verb+config.cache+ that saves the results of its tests to speed up
reconfiguring, and a file \verb+config.log+ containing compiler output
(useful mainly for debugging \verb+configure+).

@@

Another solution is to set the environment variable \verb+LD_RUN_PATH+
during linking to \verb+GLB_DIR/lib+ . Best thing is to add this to your
shell dot-file (e.g. \verb+.bashrc+). Then you can use:
A typical compiler command like
\begin{quote}
\verb+gcc -c my_program.c -IGLB_DIR/include/+
\end{quote}
and a typical linker command like
\begin{quote}
\verb#g++ my_program.o -lglobes -LGLB_DIR/lib/ -o my_executable#
\end{quote}
More information on this issue can be obtained by having a look 
into the aforementioned  \verb+Makefile+ in \verb+examples+.

\begin{itemize}
\item[CAVEAT] It is in principle possible to have many installations 
on one machine,
especially the situation of having an installation by root and by a user at
the same time might occur. However it is strictly warned against this 
possibility since it is \emph{extremely} likely to create some 
versioning problem at some time!
\end{itemize}

@@

   If you are building compiler tools for cross-compiling, you can also
use the \verb+--target=TYPE+ option to select the type of system they will
produce code for and the \verb+--build=TYPE+ option to select the type of
system on which you are compiling the package.

>>>>

The installation of \GLOBES\ is highly automated and there should not be
any problems on a decently up-to-date GNU/Linux system. The installation
has, however, only been tested on a limited number of platforms, mainly
running with various versions of SuSE Linux. Thus we would appreciate
to know your experiences with the installation on different platforms.
Please send an e-mail to <\bugs>.

\section*{Prerequisites for the installation of \GLOBES}
\index{norm}{Installation!prerequisites}

Besides the usual things such as a working libc, you need to have
\begin{itemize}
\item[gcc] The GNU compiler collection\\
\verb+gcc.gnu.org+
\item[BLAS] Basic Linear Algebra Subprograms\\
\verb+www.netlib.org/blas/+    
\item[LAPACK] Linear Algebra PACKage\\
\verb+www.netlib.org/lapack/+
\item[f2c] Fortran to C\\ 
\verb+www.netlib.org/f2c/+
\item[GSL] The GNU Scientific Library\\
\verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any
ANSI C/C++ compiler, but the \verb+globes+ binary uses the 
\verb+argp+ facility of \verb+glibc+ to parse its command line options.

All those libraries are also available as rpm's from the various
distributors of GNU/Linux: See their web sites for downloads. Due to
the rather difficult nature of the build process of BLAS and LAPACK,
it is recommended that one use a suitable rpm instead. In addition,
there is a good chance that
gcc, f2c and GSL are already part of your installation. Furthmore, you need
a working \verb+make+ to build and install \GLOBES.

In future releases, it is planned to at least make LAPACK obsolete, because
then also BLAS and f2c would drop out. Furthemore, there is no priniciple
reason why one should not get rid of any C++ parts. 
This will greatly simplify
the build process and reduce the requirements for installation.

@@

In general it is advisable to use the shared libraries. If you don't
have root privileges, see the corresponding section of this file.
If you install \GLOBES\ with root privileges, do not forget to run 
{\tt ldconfig} after installation.

@@

   The \verb+configure+ shell script attempts to guess correct values for
various system-dependent variables used during compilation.  It uses
these values to create a \verb+Makefile+ in each directory of the package.
Finally, it creates a shell script \verb+config.status+ that
you can run in the future to recreate the current configuration, a file
\verb+config.cache+ that saves the results of its tests to speed up
reconfiguring, and a file \verb+config.log+ containing compiler output
(useful mainly for debugging \verb+configure+).

@@

Another solution is to set the environment variable \verb+LD_RUN_PATH+
during linking to \verb+GLB_DIR/lib+ . Best thing is to add this to your
shell dot-file (e.g. \verb+.bashrc+). Then you can use:
A typical compiler command like
\begin{quote}
\verb+gcc -c my_program.c -IGLB_DIR/include/+
\end{quote}
and a typical linker command like
\begin{quote}
\verb#g++ my_program.o -lglobes -LGLB_DIR/lib/ -o my_executable#
\end{quote}
More information on this issue can be obtained by having a look 
into the mentioned  \verb+Makefile+ in \verb+examples+.

\begin{itemize}
\item[CAVEAT] It is in principle possible to have many installations 
on one machine. Especially the situation of having an installation by root and by a user at the same time might occur. However, it is strictly warned against this possibility, since it is \emph{extremely} likely to create some versioning problem at some time!
\end{itemize}

@@

   If you are building compiler tools for cross-compiling, you can also
use the \verb+--target=TYPE+ option to select the type of system they will
produce code for, and the \verb+--build=TYPE+ option to select the type of
system on which you are compiling the package.
